---
title:
output: github_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  echo = TRUE,
  collapse = TRUE,
  message = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/NicChr/cppdoubles/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/NicChr/cppdoubles/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/cppdoubles)](https://CRAN.R-project.org/package=cppdoubles)
[![Codecov test coverage](https://codecov.io/gh/NicChr/cppdoubles/graph/badge.svg)](https://app.codecov.io/gh/NicChr/cppdoubles)
<!-- badges: end -->

# cppdoubles

# **Fast Relative Comparisons of Floating Point Numbers in C++**

You can install `cppdoubles` using the below code.

```{r install, message = FALSE, eval = FALSE}
remotes::install_github("NicChr/cppdoubles")
```

Comparing equality of 2 double vectors

```{r basic}
library(cppdoubles)

### Basic usage ###

# Standard equality operator
sqrt(2)^2 == 2

# approximate equality operator
sqrt(2)^2 %~==% 2
```

Other approximate equality operators

```{r}
sqrt(2)^2 %~>=% 2
sqrt(2)^2 %~<=% 2
sqrt(2)^2 %~>% 2
sqrt(2)^2 %~<% 2

# Alternatively
double_equal(2, sqrt(2)^2)
double_gte(2, sqrt(2)^2)
double_lte(2, sqrt(2)^2)
double_gt(2, sqrt(2)^2)
double_lt(2, sqrt(2)^2)
```

All comparisons are vectorised and recycled

```{r}
double_equal(sqrt(1:10),
             sqrt(1:5),
             tol = c(-Inf, 1e-10, Inf))
```

One can check if a double is a whole number like so

```{r}
# One can check for whole numbers like so
whole_number <- function(x, tol = get_tolerance()){
  double_equal(x, round(x), tol = tol)
}
x <- seq(-5, 5, by = 0.2)
whole_nums <- x[whole_number(x)]
whole_nums
```

`all_equal` as an alternative to base R `all.equal.numeric`

```{r}
x <- seq(0, 10, 2)
y <- sqrt(x)^2

all_equal(x, y)
all_equal(x, 1)
all_equal(x, NA)
isTRUE(all_equal(x, NA))
```

Benchmark against `all.equal.numeric`

```{r}
library(bench)
set.seed(100)
x <- abs(rnorm(10^7))
y <- sqrt(x)^2
z <- x^2

# 2 approximately equal vectors
mean(rel_diff(x, y))
mark(base = isTRUE(all.equal(x, y)),
     cppdoubles = all_equal(x, y))

# 2 significantly different vectors
mean(rel_diff(x, z))
mark(base = isTRUE(all.equal(x, z)),
     cppdoubles = all_equal(x, z))
```


Benchmark against using absolute differences 

```{r}
mark(double_equal(x, y),
     abs_diff(x, y) < sqrt(.Machine$double.eps))
```


